---
title: "DS_EDA_Zixuan"
author: "Cherry"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readr) 
library(data.table)
library(dplyr)
library(ggplot2)
```
```{r}
products <- read.csv("products.csv")
aisles <- read.csv("aisles.csv")
departments <- read.csv("departments.csv")
orders <- read.csv("orders.csv")
train <- read.csv("order_products__train.csv")
prior <- read.csv("order_products__prior.csv")
```

```{r}
orders <- orders %>% mutate(order_hour_of_day = as.numeric(order_hour_of_day), eval_set = as.factor(eval_set))
products <- products %>% mutate(product_name = as.factor(product_name))
aisles <- aisles %>% mutate(aisle = as.factor(aisle))
departments <- departments %>% mutate(department = as.factor(department))
```

```{r}
product1 <- full_join(products,aisles,by="aisle_id")
product <- full_join(product1,departments,by="department_id")

all_order <- rbind(train,prior)
```

```{r}
orders_new <- subset(orders, select=c(order_id,user_id))
all_order_new <- subset(all_order,select = c(order_id,product_id))
m1 <- full_join(orders_new,all_order_new,by="order_id")
product_order <- full_join(product,m1,by="product_id")
# may not correct
```

#  Sales in each department(find best selling apartment)
```{r}
library(tidyr)
product_order %>%
  drop_na(department) %>%
  group_by(department) %>%
  distinct(order_id, .keep_all = T) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=department,y = count))+
  geom_bar(stat="identity",fill="red")+
  labs(title="The number of orders in each department", x="deparmtment", y="the number of orders")+
  theme(axis.text.x=element_text(angle=90, hjust=1))
  theme(plot.title = element_text(hjust = 0.5))
```
The most three popular departments are produce, dairy eggs and snacks.

# Sales in each aisle(best selling aisle)
```{r}
res <- product_order %>%
  drop_na(aisle) %>%
  group_by(aisle) %>%
  distinct(order_id, .keep_all = T) %>%
  summarise(count = n())

res <- res[c(1:15),]

res %>%
  ggplot(aes(x=aisle,y = count))+
  geom_bar(stat="identity",fill="red")+
  labs(title="The number of products in top 15 aisles ", x="deparmtment", y="the number of orders")+
  theme(axis.text.x=element_text(angle=90))+
  theme(plot.title = element_text(hjust = 0.5))
```
The top three best selling aisles are fresh fruits, fresh vegetables and packaged vegetables fruits.

# the number of reordered products
```{r,warning=FALSE}
tmp <- all_order %>% 
  group_by(reordered) %>% 
  summarize(count = n()) %>% 
  mutate(reordered = as.factor(reordered)) %>%
  mutate(proportion = count/sum(count))
kable(tmp)

tmp %>% 
  ggplot(aes(x=reordered,y=count,fill=reordered))+
  geom_bar(stat="identity")+
  labs(title="Not reorder vs Reorder", x="not reordered=0; reordered=1", y="count")+
  theme(plot.title = element_text(hjust = 0.5))
```

# most reordered product
```{r}
tmp2 <- all_order %>% 
  group_by(product_id) %>% 
  summarize(proportion_reordered = mean(reordered), n=n()) %>% 
  top_n(10,wt=proportion_reordered) %>% 
  arrange(desc(proportion_reordered)) %>% 
  left_join(product,by="product_id")
kable(tmp)

tmp2 %>%
  ggplot(aes(x=reorder(product_name,-proportion_reordered), y=proportion_reordered))+
  geom_bar(stat="identity",fill="red")+
  theme(axis.text.x=element_text(angle=90, hjust=1),axis.title.x = element_blank())+coord_cartesian(ylim=c(0.85,0.95))+
  labs(title="TOP 10 reordered rate", x="products", y="reordered rate")+
  theme(plot.title = element_text(hjust = 0.5))
```
The three products with the highest reordered rate are Raw Veggie Wrappers, Serenity Ultimate Extrema Overnight Pads and Orange Energy Shots.

# department with highest reorder ratio 
```{r}
tmp <- all_order %>%
  group_by(product_id)%>%
  summarize(proportion_reordered = mean(reordered)) %>% 
  left_join(product,by="product_id")

tmp %>%
  group_by(department) %>%
  summarize(depart_reordered = mean(proportion_reordered)) %>%
  ggplot(aes(x=department,y=depart_reordered))+
  geom_point(color="red", size=2)+
  labs(title="Reordered rate in each department", x="departments", y="reordered rate")+
  theme(axis.text.x=element_text(angle=90, hjust=1))+
  theme(plot.title = element_text(hjust = 0.5))
```
Personal care has lowest reorder ratio and dairy eggs have highest reorder ratio.

# Relationship between add_to_cart and reordered？
```{r,warning=FALSE}
tmp3 <- all_order %>%
  group_by(add_to_cart_order)%>%
  summarize(proportion_reordered = mean(reordered))

tmp3 %>%
  ggplot(aes(x=add_to_cart_order,y=proportion_reordered))+
  geom_point()+
  xlim(c(0,70))+
  ylim(c(0.3,0.7))+
  labs(title="Add to cart order - Reorder ratio", x="add to cart order", y="reordered rate")+
  theme(axis.text.x=element_text(angle=90, hjust=1))+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r,t-test}
t.test(add_to_cart_order~reordered,data=prior)
```
1.Orders placed initially in the cart are more likely to be reorderd than one placed later in the cart.
2.We did t-test to verify whether the sequence of adding to cart are siginificantly different between reordered products and not reordered products. We can conclude from the results showing the p-value is smaller than 0.05 that the sequence of adding to cart significantly influence whether the products being reordered.