---
title: "DS_EDA_Ruth"
author: "Cherry"
date: "11/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readr) 
library(dplyr)
library(ggplot2)
library(corrplot)
library(knitr)
library(tidyr)
products <- read.csv("~/Downloads/instacart-market-basket-analysis/products.csv")
aisles <- read.csv("~/Downloads/instacart-market-basket-analysis/aisles.csv")
departments <- read.csv("~/Downloads/instacart-market-basket-analysis/departments.csv")

product1 <- full_join(products,aisles,by="aisle_id")
finalproduct <- full_join(product1,departments,by="department_id")
```


### How many products/aisles/departments?
```{r}
nrow(aisles)
nrow(departments)
nrow(products)
str(finalproduct)
```
* There are `r nrow(aisles)` aisles in the data
* There are `r nrow(departments)` departments in the data
* There are `r nrow(products)` products in the data

### Check the 'finalproduct' dataset(eg:head,structure,missing value)
```{r}
glimpse(finalproduct)
head(finalproduct)
sum(complete.cases(finalproduct))
```
There are 'r sum(complete.cases(finalproduct))` complete rows in the data, hence no missing observation

### How many products in each department?
```{r}
prod_by_dept <- finalproduct %>%
                                      group_by(department)%>%
                                       tally()
prod_by_dept
color_range <- colors()
ggplot(prod_by_dept,aes(reorder(department,-n),n))+geom_bar(stat = "identity", fill= 'blueviolet')+
                                      theme(axis.text.x=element_text(angle=45, hjust=1), axis.title.x = element_blank())
```


### Top 5 departments with the most products?
```{r}
dept_top_five <- finalproduct %>%
                                       group_by(department)%>%
                                       tally()%>%
                                       arrange(desc(n))%>%
                                       top_n(5,n)

head(dept_top_five)

#install.packages("plotrix")
library(plotrix)
pie3D(dept_top_five$n, labels = dept_top_five$department, main = "An exploded 3D pie chart of top 5 Departments", explode=0.1, radius=.9, labelcex = 1.2,  start=0.7)
```


### How many products in each aisle?
```{r}
prod_by_aisle <- finalproduct %>% group_by(aisle)%>% tally()%>%
                                       arrange(desc(n))
                                       
prod_by_aisle
```

###  Top 10 aisles with the most products?
```{r}
aisle_top_ten <- finalproduct %>%
                                       group_by(aisle)%>%
                                       tally()%>%
                                       arrange(desc(n))%>%
                                       top_n(10,n)
head(aisle_top_ten)
ggplot(aisle_top_ten,aes(reorder(aisle,-n),n))+geom_bar(stat = "identity", fill='coral2')+
                                      theme(axis.text.x=element_text(angle=45, hjust=1),axis.title.x = element_blank())
```


